A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-01-29 archive/v3/docs_v3/discussion_merger_v3.04.tex
Cleaned up (files are moved to archive now)

\documentclass[useAMS,usenatbib]{mnras}

%_____GENERAL PACKAGES_____
\usepackage{amsmath}	%for \text{} in math mode
\usepackage{amssymb}	%for symbols

%_____FIGURES_____
\usepackage{graphicx}	%for including images
\usepackage{float}	%for [Hhtb] etc of figures
\usepackage{grffile}    %for using underscores in figure filenames
\usepackage{subcaption} %for having sub-captions in figures with multiple panels
\usepackage{placeins} %for \FloatBarrier, which forces all figures to be printed before continuing

%_____SYMBOLS_____
\usepackage{ mathrsfs } %for slanted letters in math mode

%_____FORMAT_____
\usepackage{hyperref} %ready-to-lick referencing

%_____MACROS______
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
\newcommand{\vect}[1]{\vec{#1}} %  ARROW vectors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[A spiral galaxy's mass distribution uncovered]{A spiral galaxy's mass distribution uncovered through lensing and dynamics}
\author[W.H. Trick, G. van de Ven and A.A. Dutton]{Wilma H. Trick$^{1}$\thanks{E-mail:
trick@mpia.de}, Glenn van de Ven$^{1}$ and Aaron A. Dutton$^{1}$\\
$^{1}$Max-Planck-Institute for Astronomy, K\"{o}nigstuhl 17, 69117 Heidelberg, Germany}
\begin{document}

\date{Accepted ???. Received ???; in original form ???}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

\maketitle

\label{firstpage}

%--------------------------------------------
\begin{abstract}
We investigate the matter distribution of a spiral galaxy with counter-rotating stellar core, SDSS J1331+3628 (J1331), independently with gravitational lensing and stellar dynamical modelling. By fitting a gravitational potential model to a quadruplet of lensing images around J1331's bulge, we tightly constrain the mass inside the Einstein radius $R_\text{ein}=(0.91\pm0.02)''(\simeq1.83\pm0.04~\text{kpc})$ to within 4\%: $M_\text{ein} = (7.8\pm0.3) \cdot 10^{10} M_\odot$. We model observed long-slit major axis stellar kinematics in J1331's central regions by finding Multi-Gaussian Expansion (MGE) models for the stellar and dark matter distribution that solve the axisymmetric Jeans equations. The lens and dynamical model are independently derived, but in very good agreement with each other around $\sim R_\text{ein}$. We find that J1331's center requires a steep total mass-to-light ratio gradient. A dynamical model including a NFW halo (with virial velocity $v_{200} \simeq 240 \pm 40~\text{km s}^{-1}$ and concentration $c_{200} \simeq 8 \pm 2$) and moderate tangential velocity anisotropy ($\beta_z \simeq -0.4 \pm 0.1$) can reproduce the signatures of J1331's counter-rotating core and predict the stellar and gas rotation curve at larger radii. However, our models do not agree with the observed velocity dispersion at large radii. We speculate that the reason could be a non-trivial change in structure and kinematics due to a possible merger event in J1331's recent past.
\end{abstract}
%--------------------------------------------

\begin{keywords}
gravitational lensing: strong -- stars: kinematics and dynamics -- galaxies: kinematics and dynamics -- galaxies: photometry -- galaxies: structure
\end{keywords}

%--------------------------------------------
\section{Introduction}
%--------------------------------------------

%--------------------------------------------
Gravitational lensing and dynamical modelling provide independent constraints on the mass distribution of galaxies. Combining them allows for valuable cross-checking opportunities to disentangle the galactic stellar and dark matter (DM) content.

Cosmological simulations suggest that cold DM forms cuspy halos following a Navarro-Frenk-White (NFW) profile \citep{1996ApJ...462..563N}. However, the existence of central DM density cusps in massive galaxies depends strongly on the stellar mass-to-light ratio (e.g., \citealt{2011MNRAS.416..322D}) and DM dominated dwarf galaxies even favour DM halos with cores (e.g., \citealt{1994Natur.370..629M,2001ApJ...552L..23D}). This discrepancy, known as the core-cusp problem, might be resolved by galaxy formation processes such as mergers and outflows (e.g., \citealt{2001ApJ...560..636E,2012MNRAS.421.3464P}).

Determining the overall mass distribution in massive galaxies and separating the DM from the stellar mass components is therefore a crucial step in better understanding the structure and formation of galaxies and nature of DM.

Massive galaxies can act as gravitational lenses, deflect the light of background sources and give rise to multiple images. This strong gravitational lensing tightly constrains the projected total mass of the lens galaxy inside $\sim 1''$ (e.g., \citealt{2010ARA&A..48...87T}).

The mass profile at larger galactocentric radii can be probed by gas rotation curves that directly measure the galaxy's circular velocity profile (e.g., \citealt{1980ApJ...238..471R}). However, due to its dissipative nature gas motions are very sensitive to disturbances by, e.g., spiral arms and bars (e.g., \citealt{2004dad..book.....S}).

Because stars are dissipationless dynamical tracers and present almost everywhere in the galaxy, stellar dynamical modelling can complement mass constraints from lensing at small and gas motions at large radii. As stellar motions are complex---a bulk rotation around one principal axis combined with random motions in all coordinate directions---\citep{2008gady.book.....B}, full dynamical modelling of rotation, dispersion and velocity anisotropies is needed to deduce the matter distribution.

The Sloan WFC Edge-on Late-type Lens Survey (SWELLS, WFC = Wide field camera) \citep{SWELLSI,SWELLSII,SWELLSIII,SWELLSIV,SWELLSV,SWELLSVI} is dedicated to find and investigate spiral galaxies, which are (i) strong gravitational lenses and (ii) observed almost edge-on, such that rotation curves can be easily measured. By combining lensing and dynamical modelling, degeneracies inherent in both methods can be broken.

One of the SWELLS galaxies is the massive spiral SDSS J1331+3628, to which we refer as J1331 in the remainder of this work. It has bluish spiral arms and a large reddish bulge (see Figures \ref{fig:F450W} and \ref{fig:F814W}), which is superimposed by a quadruplet of extended bluish images approximately at a distance of $1''$ from the galaxy center (see Figure \ref{fig:lens_just_imgpos}). The lensed object might be a star-forming blob of a background galaxy. J1331 stands out of the SWELLS sample because of its large counter-rotating core (see Figure \ref{fig:kinematics}), which might be an indication that J1331 underwent a merger in its recent past.

\citet{SWELLSI} confirmed that J1331 is a strong gravitational lens, measured its apparent brightness and estimated the stellar masses of disk and bulge. The lensing properties of J1331 were first analysed by \citet{SWELLSIII}. \citet{SWELLSV} measured the gas and stellar kinematics along the major axis and deduced J1331's mass profile from the gas rotation curve at large radii and total mass inside the Einstein radius from gravitational lensing, focusing mostly on the outer regions of J1331.

The goal of this work is to complement the previous studies of J1331 by an in-depth analysis of the matter distribution in J1331's inner regions. We use stellar dynamical modelling in addition to lensing constraints, similar to a study of the Einstein Cross by \citet{GlennEC}. We attempt to disentangle the stellar and DM components and test if employed axisymmetric Jeans models work also in the presence of a counter-rotating core. Ideally, this work on J1331 could also help understanding how mergers modify the mass distribution of a galaxy.

This paper is organized as follows: Section \ref{sec:data} summarizes the data, Section \ref{sec:Modelling} gives an overview of the modelling techniques used in this work, and Section \ref{sec:Results} presents our results on the surface photometry of J1331 using Multi-Gaussian expansions (Section \ref{sec:MGE_results}), constraints from lensing (Section \ref{sec:results_lensing}) and Jeans modelling based on the surface brightness only (Section \ref{sec:results_JAM_SB}) and including a NFW DM halo (Section \ref{sec:results_JAM_NFW}). Finally, Section \ref{sec:Discussion} uses these results to discuss J1331's stellar mass-to-light ratio, central kinematics, possible merger history and starting points for future work.
%--------------------------------------------

%============================================
\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in the F450W filter.}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in the F814W filter.}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{Lensing images around the galaxy center.}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar kinematics by \citet{SWELLSV}.}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space Telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3628 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The black solid line in panel (b) shows the orientation of the major axis. Its length is $10''$ and it marks approximately the region within which we carry out the dynamical modelling. \emph{Panel (c):} Lensing images in the central region of J1331.  An IRAF \emph{ellipse} model was fitted to and then subtracted from the galaxy's F450W surface brightness in panel (a). The (smoothed) residuals within the white square in panel (a) are shown in panel (c). The four bright blobs (A, B, C and D), which are visible in the residuals, are arranged in a typical strong lensing configuration around the center of the galaxy (G). (The configuration of the two additional blobs which lie approximately on a line with A, B and G does not suggest that these blobs are a lensing doublet. They might rather be star-forming regions of a background galaxy.) \emph{Panel (d):} Stellar kinematics along the galaxy's major axis as measured by \citet{SWELLSV}. Shown are line-of-sight rotation velocity $v_\text{rot}$ (top), line-of-sight velocity dispersion $\sigma$ (middle) and the root mean-square (rms) velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$ (bottom). The dashed line indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6'' \hat{=} 5.2~\text{kpc}$. The $v_\text{rot}$ curve reveals that J1331 has a counter-rotating core within $R_\text{eff}$.}
\label{fig:specialJ1331}
\end{figure*}
%============================================

%--------------------------------------------
\section{Data} \label{sec:data}
%--------------------------------------------

%--------------------------------------------
\subsection{Redshift and position}
%--------------------------------------------

%--------------------------------------------
J1331 is located at right ascension = 202.91800$^\circ$ and declination = 36.46999$^\circ$ (epoch J2000). \citet{SWELLSI} found from SDSS spectra that J1331 has two redshifts inside $R'=1''$: J1331 itself has $z_d = 0.113$ and $z_s = 0.254$ is the redshift of the lensed background source \citep{SWELLSIII}. According to the WMAP5 cosmology \citep{WMAP5cosm}, J1331 has an angular diameter distance of 414 Mpc, which translates into a transverse scaling of $1'' \hat{=} 2.01~\text{kpc}$.

%--------------------------------------------
\subsection{HST imaging}
%--------------------------------------------

%--------------------------------------------
We use HST imaging of J1331 by \citet{SWELLSI}. They performed high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) and its WF3 CCD chip. The images are a combination of four exposures with each an exposure time of $400~\text{sec}$ and were drizzled to a pixel scale of 1 pixel = $0.05''$. In particular, we use the images in the filters F450W, to identify the positions of the bluish lensing images, and F814W (I-band) to create a surface brightness model of the reddish bulge.

%--------------------------------------------
\subsection{Stellar kinematics} \label{sec:data_kinematics}
%--------------------------------------------

%--------------------------------------------
For the dynamical modelling we use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slit was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. 1D spectra for spatial bins of different widths along the major axis were extracted. Line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersions ($\sigma$) were determined from the spectral region containing the absorption lines Mgb (5177 \AA) and FeII (5270, 5335 and 5406 \AA) (analogous to \citealt{SWELLSII}). Gas kinematics were extracted from Gaussian line profile fits to H$\alpha$ (6563 \AA) and [NII] (6583 and 6548 \AA) emission lines as tracers for ionized gas.

The stellar kinematics, $v_\text{rot}$, $\sigma$ and the root-mean-square velocity $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\hat{=}4~\text{kpc}$. Outside of $\sim 3.5''$ there is a drop in the dispersion. This could indicate the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W} the smooth reddish bulge extends out to $\sim5''$.

Inside of $\sim 4''$ the data appears to be symmetric; outside of it the data suggests that the assumption of axisymmetry seems not to be valid anymore. We subtract $2.3~\text{km s}^{-1}$ from $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} \geq 5~\text{km s}^{-1}$ to the $v_\text{rms}$ data. In our dynamical Jeans modelling approach, which is based on the assumption of axisymmetry, only stellar kinematics with either $R'  \lesssim 3.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict the modelling to the bulge region is that our surface brightness model (Table \ref{tab:MGEF814W} and Figure \ref{fig:MGEinnerRegions}) is only a good representation of J1331's F814W light distribution inside $\sim 5''$.
%--------------------------------------------

%--------------------------------------------
\section{Modelling} \label{sec:Modelling}
%--------------------------------------------

%--------------------------------------------
\subsection{Surface brightness model} \label{sec:MGE_theo}
%--------------------------------------------

%--------------------------------------------
\subsubsection{Multi-Gaussian Expansion (MGE)}
%--------------------------------------------

%--------------------------------------------
MGEs are used to parametrize the observed surface brightness (or projected total mass) of a galaxy as a sum of $N$ 2-dimensional, elliptical Gaussians \citep{1991ApJ...366..599B,1992A&A...253..366M,1994A&A...285..723E,1999MNRAS.303..495E}. This work makes use of the algorithm and code\footnote{The IDL code package for fitting MGEs to images by \citet{Cap02} is available online at \url{http://www-astro.physics.ox.ac.uk/~mxc/software}. The version from June 2012 was used in this work.} by \citet{Cap02}. We assume all Gaussians to have the same center and position angle $\phi$, i.e., the orientation of the Gaussians' major axis measured from North through East in the polar coordinate system $(R',\theta')$ on the plane of the sky. Then a surface brightness model can be expressed as
\begin{eqnarray}
I(R',\theta') &=& \sum_{i=1}^{N} I_{0,i} \exp\left[ - \frac{1}{2\sigma_i^2} \left(x^{'2} + \frac{y^{'2}}{q_i^{'2}}\right)\right]\label{eq:MGEgeneral}\\
\text{with } I_{0,i} &=& \frac{L_i}{2\pi \sigma_i^2 q'_i}\label{eq:centralItotalL}\\
\text{and } x'_i &=& R' \cos(\theta' - \phi)\nonumber\\
y'_i &=& R' \sin(\theta' - \phi),\nonumber
\end{eqnarray}
where $I_{0,i}$ is the central surface brightness of each Gaussian, $L_i$ its total luminosity, $\sigma_i$ its dispersion along the major axis and $q'_i$ the axis ratio between the elliptical Gaussian's major and minor axis.

\subsubsection{Convolution with the point spread function (PSF)}

We can also expand the telescope's PSF as a sum of circular Gaussians,
\begin{equation}
\text{PSF}(x',y') = \sum_j \frac{G_j}{2 \pi \delta_j^2} \exp\left[- \frac{1}{2 \delta_j^2} \left({x'}^2 + {y'}^2 \right)\right], \label{eq:PSFgeneral}
\end{equation}
where $\sum_j G_j = 1$ and $\delta_j$ are the dispersions of the circular PSF Gaussians.

The observed surface brightness distribution is a convolution of the intrinsic surface brightness in Equation \eqref{eq:MGEgeneral} with the PSF in Equation \eqref{eq:PSFgeneral}: $(I \ast \text{PSF}) (x',y')$ is then again a sum of Gaussians and can be directly fitted to an image of the galaxy in question.

\subsubsection{Deprojection} \label{sec:MGE_theo_deprojection}

$I(R',\theta')$ describes the intrinsic, to 2D projected light distribution or surface density of the galaxy. Under the assumption that the galaxy is oblate and axisymmetric, and given the inclination angle $i$ of the galaxy with respect to the observer, MGEs allow an analytic deprojection of the 2-dimensional MGE to get a 3-dimensional axisymmetric light distribution or density $\nu(R,z)$ for the galaxy,
\begin{equation}
\nu(R,z) = \sum_i \nu_{0,i} \exp \left[-\frac{1}{2\sigma_i}\left(R^2 + \frac{z^2}{q_i^2} \right) \right], \label{eq:deprojMGE}
\end{equation}
where $R$ is the distance from the galaxy's short axis and $z$ the height above the galactic plane. The flattening $q_i$ of each axisymmetric 3D Gaussian and its central density $\nu_{0,i}$ follow from the observed 2D axis ratio $q'_i$ and surface density $I_{0,i}$ as
\begin{eqnarray*}
q_i^2 &=& \frac{q_i'^2 - \cos^2 i}{\sin^2 i}\\
\nu_{0,i} &=& \frac{q_i' I_{0,i}}{q_i \sqrt{2 \pi \sigma_i^2}}.
\end{eqnarray*}
%--------------------------------------------

%--------------------------------------------
\subsection{Strong gravitational lens model} \label{sec:lensing_theo}
%--------------------------------------------

%--------------------------------------------
\subsubsection{Strong lensing formalism}
%--------------------------------------------

%--------------------------------------------
A gravitational lens is a mass distribution whose gravitational potential $\Phi$ acts as a lens for light coming from a source positioned somewhere on a plane behind the lens. We summarize the basic formalism (see also \citealt{1992grle.book.....S,1996astro.ph..6001N,2006glsw.conf....1S,2006glsw.conf...91K,2010ARA&A..48...87T}). The angular diameter distance from the observer to the lens is $D_d$, to the source plane $D_s$, and the distance between the lens and source plane is $D_{ds}$. The deflection potential of the lens is its potential projected along the line of sight $z'$ and rescaled to
\begin{equation}
\psi(\vect{\theta}) := \frac{D_{ds}}{D_d D_s} \frac{2}{c^2} \int \Phi(\vect{r}=D_d \vect{\theta},z') {\ \mathrm d} z', \label{eq:psidef}
\end{equation}
where $\vect{\theta}\equiv(x',y')$ is a 2-dimensional vector on the plane of the sky. The light from the source at $\vect{\beta} \equiv (x_s',y_s')$ is deflected according to the lens equation
\begin{equation}
\vect{\beta} = \vect{\theta}_i - \left.\vect{\nabla}_\theta \psi(\vect{\theta})\right|_{\vect{\theta}_i} \label{eq:lenseqpot}
\end{equation}
into an image $\vect{\theta}_i = (x_i',y_i')$. The gradient of the deflection potential, $\vect{\nabla}_\theta \psi(\vect{\theta})$, is equal to the angle by which the light is deflected multiplied by $D_{ds}/D_{s}$.

The total time delay of a deflected light path through $\vect{\theta}$ with respect to the unperturbed light path is given by 
\begin{equation}
\Delta t(\vect{\theta}) = \frac{(1+z_d)}{c} \frac{D_d D_s}{D_{ds}} \left[ \frac 12 (\vect{\theta} - \vect{\beta})^2 - \psi(\vect{\theta})\right]. \label{eq:timedelay}
\end{equation}
 According to Fermat's principle the image positions will be observed at the extrema of $\Delta t(\vect{\theta})$.

The inverse magnification tensor
\begin{equation}
\mathscr{M}^{-1} \equiv \frac{\partial \vect{\beta}}{\partial \vect{\theta}} \overset{(\ref{eq:lenseqpot})}{=} \left(\delta_{ij} - \frac{\partial^2 \psi}{\partial \theta_i \partial \theta_j} \right)\label{eq:magnificationtensor}
\end{equation}
describes how the source position changes with image position. It also describes the distortion of the image shape for an extended source and its magnification $\mu$ due to lensing according to
$$\mu \equiv \frac{\text{image area}}{\text{source area}} = \det \mathscr{M}.$$
Lines in the image plane for which the magnification becomes infinite, i.e., $\det \mathscr{M}^{-1} = 0$, are called critical curves. The corresponding lines in the source plane are called caustics. The position of the source with respect to the caustic detemines the number of images and their configuration and shape with respect to each other.

The Einstein mass $M_\text{ein}$ and Einstein radius $R_\text{ein}$ are defined via the relation
\begin{equation*}
M_\text{ein} \equiv M(<R_\text{ein}) \overset{!}{=} \pi \Sigma_\text{crit} R_\text{ein}^2,
\end{equation*}
where $\Sigma_\text{crit} \equiv \frac{c^2}{4\pi G} \frac{D_s}{D_d D_{ds}}$ is the critical density and $M(<R_\text{ein})$ is the mass projected along the line-of-sight within $R'=R_\text{ein}$. $M_\text{ein}$ is similar to the projected mass within the critical curve $M_\text{crit}$.

%--------------------------------------------
\subsubsection{Lens model} 
%--------------------------------------------

%--------------------------------------------
Following \citet{EvansWitt} we assume a scale-free model
\begin{equation}
\psi(R',\theta') = R^{'\alpha} F(\theta') \label{eq:scalefreemodel}
\end{equation}
for the lensing potential, consisting of an angular part $F(\theta')$ and a radial power-law part, with $(R',\theta')$ being again polar coordinates on the plane of the sky. $\alpha$ denotes the rotation curve slope of the lens and the case $\alpha = 1$ corresponds to a flat rotation curve. The surface density of this model is 
\begin{equation*}
\Sigma(R',\theta')= \frac{ \Sigma_\text{crit}}{2} \left(\alpha^2 F(\theta') + \frac{\partial^2}{\partial \theta^{'2}}F(\theta')\right) R^{'\alpha-2}.
\end{equation*}
We expand $F(\theta')$ into a Fourier series,
\begin{equation}
F(\theta') = \frac{a_0}{2} + \sum_{k=1}^{\infty} \left(a_k \cos(k\theta') + b_k \sin (k\theta') \right). \label{eq:Fourieransatz}
\end{equation}
For this scale-free lens model the lens equation \eqref{eq:lenseqpot} becomes
\begin{equation}
\begin{pmatrix} x_s' \\ y_s' \end{pmatrix} = \begin{pmatrix} R'_i \cos \theta'_i - R_i^{'\alpha-1} \left(\alpha \cos \theta'_i F(\theta'_i) - \sin \theta'_i F'(\theta'_i) \right) \\ R'_i \sin \theta'_i - R_i^{'\alpha-1} \left(\alpha \sin \theta'_i F(\theta'_i + \cos \theta'_i F'(\theta'_i) \right)\end{pmatrix}\label{eq:Fourierlenseq}
\end{equation}
\citep{EvansWitt}, where $F'(\theta') \equiv \partial F(\theta') / \partial \theta'$. When we fix the slope $\alpha$, then the lens equation is a purely linear problem and can be solved numerically for the source position $(x_s',y_s')$ and the Fourier parameters $(a_k,b_k)$ given one observed image at position $(x'_i=R'_i \cos \theta'_i,y'_i=R'_i \sin \theta'_i)$.

%--------------------------------------------
\subsubsection{Model fitting}
%--------------------------------------------

%--------------------------------------------
The free parameters of the lens model are: the source position $(x_s',y_s')$, the radial slope $\alpha$ and Fourier parameters $(a_k,b_k)$ of the lens mass distribution in Equations \eqref{eq:scalefreemodel}-\eqref{eq:Fourieransatz}. We want to minimize the distance between the observed image positions, $\vect{\theta}_{oi}$, and those predicted by the lensing model, $\vect{\theta}_{pi}$. To avoid having to solve the lens equation \eqref{eq:Fourierlenseq} for $\vect{\theta}_{pi}$, we follow \citet{1991ApJ...373..354K} and cast the calculation back to the source plane using the magnification tensor in Equation \eqref{eq:magnificationtensor} to approximate $\vect{\theta} \simeq (\partial \vect{\theta} / \partial \vect{\beta}) \vect{\beta} = \mathscr{M} \vect{\beta} $. The best fit lens model is then the one that minimizes
\begin{eqnarray}
\chi^2_\text{lens} &=& \sum_{i} \left|\left( \begin{matrix} \frac{1}{\Delta_x} & 0\\0 & \frac{1}{\Delta_y}\end{matrix}\right) \left( \vect{\theta}_{pi} - \vect{\theta}_{oi} \right)\right|^2\nonumber\\
&\simeq& \sum_{i} \left|\left( \begin{matrix} \frac{1}{\Delta_x} & 0\\0 & \frac{1}{\Delta_y}\end{matrix}\right)  \left.\mathscr{M}\right|_{\vect{\theta}=\vect{\theta}_{oi}} \left( \begin{matrix} x_s' - \tilde{x}'_{si} \\ y_s' - \tilde{y}'_{si} \end{matrix} \right) \right|^2,\label{eq:chi2lens}
\end{eqnarray}
where $(\Delta_x,\Delta_y)$ are the measurement errors of the image positions $\vect{\theta}_{oi}$. $\left.\mathscr{M}\right|_{\vect{\theta}=\vect{\theta}_{oi}}$ is the magnification tensor and $(\tilde{x}'_{si},\tilde{y}'_{si})$ the source position according to the lens equation, both evaluated at the position of the $i$-th lensing image, $\vect{\theta}_{oi}$. Following \citet{GlennEC} we add a term
\begin{equation}
\chi^2_\text{shape} = \lambda \sum_{k \geq 3} \frac{\left(a_k^2 +b_k^2 \right)}{a_0^2} \label{eq:chi2shape}
\end{equation}
(with $\lambda$ being some weight factor) which forces the shape of the mass distribution to be close to an ellipse. The total $\chi^2$ to minimize is therefore
\begin{equation}
\chi^2 = \chi^2_\text{lens} + \chi^2_\text{shape} \label{eq:chi2total}
\end{equation}
We set $a_1 = b_1 = 0$, which corresponds to the choice of origin; in this case the center of the galaxy.

To be able to constrain the slope $\alpha$ we would need flux ratios for the images as in \citet{GlennEC}. But the extent of the images, possible dust obscuration and surface brightness fluctuations due to microlensing events, as well as the uncertainty in surface brightness subtraction, make flux determination too unreliable and we do not include them in the fitting.
%--------------------------------------------

%--------------------------------------------
\subsection{Axisymmetric dynamical model} \label{sec:model_JAM}
%--------------------------------------------

%--------------------------------------------
\subsubsection{Jeans Anisotropic Models (JAM)} \label{sec:model_JAM_JAM}
%--------------------------------------------

%--------------------------------------------
JAM modelling assumes galaxies to be (i) collisionless (i.e., the collisionless Boltzmann equation for the distribution function $f(\vect{x},\vect{v},t)$ has to be satisfied, $\frac{\diff f(\vect{x},\vect{v},t)}{\diff t} = 0$), (ii) in a steady state ($\frac{\partial}{\partial t} = 0$), (iii) axisymmetric (best described in cylindrical coordinates $(R,z,\phi)$ and $\frac{\partial}{\partial \phi} = 0$). From this follow the axisymmetric Jeans equations as the vector-valued first moment of the Boltzmann equation, i.e.,
\begin{equation*}
\int \vect{v} \frac{\diff f}{\diff t} \Diff3 v = 0.
\end{equation*}
To be able to solve the Jeans equations, additional assumptions about the velocity ellipsoid tensor $\langle v_i v_j\rangle$ have to be made. We follow \citet{Cap08} and assume firstly that the galaxy's velocity ellipsoid is aligned with the cylindrical coordinate system, i.e., $\langle v_R v_z\rangle = \langle v_z v_\phi \rangle = \langle v_\phi v_R \rangle = 0$. Secondly, we assume a constant ratio between the radial and vertical second velocity moments, 
\begin{equation}
\beta_z \equiv 1 - \langle v_z^2 \rangle / \langle v_R^2\rangle, \label{eq:bz}
\end{equation}
where $\beta_z$ is the velocity anisotropy parameter. This reduces the Jeans equations to two equations for $\langle v_z^2 \rangle$ and $\langle v_\phi^2 \rangle$ which can be solved by means of one integration,
\begin{eqnarray}
n \langle v_z^2 \rangle (R,z) &=& \int_0^\infty n \frac{\partial \Phi}{\partial z} \diff z \label{eq:Jeans1}\\
n \langle v_\phi^2 \rangle (R,z) &=& R \frac{\partial}{\partial R} \left( \frac{n \langle v_z^2 \rangle}{1-\beta_z} \right) + \frac{n \langle v_z^2 \rangle}{1 - \beta_z} + R n \frac{\partial \Phi}{\partial R},\label{eq:Jeans2}
\end{eqnarray}
where $n(\vect{x}) \equiv \int f(\vect{x},\vect{v}) \Diff3 v$ is the number density of (stellar) tracers and $\Phi(\vect{x})$ the galaxy's gravitational potential generated by the mass density $\rho(\vect{x})$.

>>>>

\subsection{On J1331's possible merger history and modelling failures}

%Why we think that J1331 had a merger
J1331 has a large counter-rotating stellar core within $\sim 2''$. This suggests a process in J1331's past in which two components with angular momenta oriented in opposite directions were involved.

%Gas accretion - no!
Accretion of gas on retrograde orbits and subsequent star formation could lead to a younger and counter-rotating stellar population \Wilma{[TO DO: REF]}. However, to form enough stars such that the net rotation of the large and massive core is retrograde, a very large amount of gas would have had to be accreted by J1331---which is not very likely.

%Mergers - yes!
Another scenario are galaxy mergers. Major mergers can form kinematically decoupled cores (KDCs) (e.g., \citealt{2011MNRAS.414.2923K,2015ApJ...802L...3T}), if they include large amounts of gas \citep{2010ApJ...723..818H}. During a minor merger, the dense nucleus of a satellite galaxy on a retrograde orbit could survive the dissipationless accretion and spiral to the core due to tidal friction (e.g., \citealt{1984ApJ...287..577K}).

%Merger + color gradient
Usually ellipticals and the bulges of massive spirals appear reddest in their center and getting increasingly bluer with larger radii. Mergers can reverse this behaviour by inducing the creation of young stellar populations in the remnant's center. Major mergers can trigger star formation bursts (e.g., \Wilma{[TO DO: REF, maybe http://arxiv.org/pdf/1409.5126v2.pdf???]}). After a minor merger the satellite's stellar population now residing in the remnant's core is in general younger than the progenitor's bulge \Wilma{[TO DO: REF]} . The different stellar populations in a merger remnant can be associated with different $\Upsilon_*$ and sometimes even show up as a reverse colour gradient within the bulge in photometry \Wilma{[TO DO: REF]}.

%Color gradient in J1331
Even though investigation of the photometry of J1331 did not reveal a distinct blue core in J1331, we cannot fully exclude the possibility that J1331 has such a $\Upsilon_*$ gradient (see discussion in Section  \Wilma{[TO DO: Where???]}). In this case the restrictive assumption of a constant $\Upsilon_*$ in our JAM modelling could have lead to wrong conclusions about the DM distribution.

%Merger and kinematic twist
Other ways how mergers can can modify the structure of galaxies are warps in the outer parts or kinematic twists in the center of the remnant galaxy \citep{2013pss5.book..923S}, which show up as misalignments of kinematic and photometric major axis.

%Kinematic twist in J1331
From kinematics along the photometric major axis only it cannot be determined if such a misalignement is present in J1331. But if it were, our modelling assumption of axisymmetry would not be satisfied anymore.

We note that kinematics measured along an axis different from the kinematic major axis would result in $v_\text{rot}$ measurements being lower than the actual maximum $v_\text{circ}$. At large radii this could lead to an underestimation of the DM halo mass.
